|

### 2021-09-05 appendix/appendix.Rmd
changes paper

```{r, include=FALSE}
## Starting
#XXXX

# Required packages
#XXXX

# Load datasets
#XXXX
```

@@

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors \textbf{x}.

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure majority of council votes. The decision process leaves the mayor's office, and either immediately stops or it is transferred to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the council in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan affiliation; a \textit{partisan} reversal mechanism, where the council is divided in supporters and non-supporters of the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan with non-partisan features. In the main paper, we focus on the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that the partisan a mechanism is what drives our results.

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals. Similar to \citet{buenodemesquita2005logic}, we assume that the selectorate, i.e. the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, which is the minimum number of members that have to be convinced to implement the policy, is equal to half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor or the councilor recognized as the proposer \textbf{x}. We leave the definition of these transfers open, as this facilitates the relation with several instruments that mayors can use to pressure city councilors. To illustrate, examples of transfers are granting electoral favors, spending in areas that the councilor has political or personal interests, granting portfolio and public jobs within the municipal bureaucracy, or even campaign contributions and bribes. The use of transfers to affect the councilors' choices and generate a governing cost of $C_G$. The governing costs vary with the reversal mechanism and the legislature size.

Finally, after paying the costs of govern, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. We assume that the benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to say that mayors have a strong preference to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R > 0$ resources, and cannot run debts, meaning that the budget has to be balanced. We also assume that there are enough resources for the mayor to govern.

Let $C_G$ be the costs of govern. The budget balance constraint requires the offers that the mayor makes to the city councilors satisfy the following inequality:

@@

The expected utility for the city councilors depend on the reversal policy. As discussed, we explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process repeats, until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, we say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O  = \{1, 2, \cdots, N \} \setminus G$. The mayor's policy generate a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends of the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor, that would receive $p$ from the policy, now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy, and now receives zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in terms of public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent to defy the partisan label.

In the last reversal mechanism, the \textit{hybrid} reversal mechanism, we combine partisanship with legislature bargaining rents. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the non-partisan reversal mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. Opposition members, on the other hand, receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as city councilors do not control the Executive when the two branches are separated, no public goods are provided ($p=0$). Moreover, in equilibrium, we assume that city councilors would prefer to receive rents instead of providing public goods. This aligns with the empirical evidence in most developing economies.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, we have three different reversal policies. The details are in the following sections.
\end{enumerate}

This is a bargaining game, along the lines of \citet{rubinstein1982perfect}. For a political science models similar to this model, see \citet{baron1989bargaining} and, to some extent, \citet{buenodemesquita2005logic}.

The solution concept we use in this model is the Sub-game Perfect Nash Equilibrium. A Sub-game Perfect Nash Equilibrium requires that the strategies played are a Nash Equilibrium in each of the subgames of the game. The solution is found using backward induction. In the model with infinitely repeated proposals, we extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium is based on the assumption that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should accept the same offer at time $k$. Stationarity gives us an objective of finding an offer that would be accepted at any stage in the game. Applying this reasoning, we can find a sequence of offers for each point $k$ in time. For the mayor the optimal offer is at $k=0$, and this represents a no delay in policy implementation.

\begin{definition}
The equilibrium for the game consists in a sequence of strategic policy vectors $(g^*, r^*, x^*)_k$, indexed in $k \in \{0, 1, \cdots, \infty\}$, such that, at every stage of the game, the mayor and the councilors have no profitable deviations from the equilibrium.
\end{definition}

Moreover, we assume that the players play no dominated strategies, even when there is less than sufficient support for implement the policy.

@@

Solving the game by backward induction would require us to start with the councilor's strategies. However, for simplicity, assume at this point that we solved the the game for the councilors, and found the equilibrium cost of govern $C_G(N)$ when there are $N$ city councilors. Below we define $C_G(N)$ for each of the mechanisms, but for now we use this assumption to simplify the exposition.

We now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as necessary for the public goods, and extract the remaining resources as political rents. The mayor's objective is to maximize the following expected utility, subject to the municipal budget constraint.

@@

In equilibrium the budget binds, and the public goods provision maximizes:

@@

The first order condition for an optimal public goods provision makes the marginal benefits to increase rents, in this case, marginal costs of providing the public goods, equals to the marginal benefits of reelection:

@@

\begin{proposition}
The provision of public goods increase with legislature size when the costs of govern decrease with legislature size.
\end{proposition}

\begin{proof}
To prove this result we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a the cross-partial derivative on $N$ or even use the implicit function theorem. Instead, we use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

The mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1 > N$,\footnote{Moreover, the requirement that the decision and type space is a complete lattice, is satisfied for $\mathbb{R} \times \mathbb{N}$.} we have:

@@

After substituting, we have:

@@

Multiplying both sides by $\dfrac{1}{g - g'}$ and taking the limit when $g \rightarrow g'$, we have:

@@

And the function satisfies the increasing differences in $g$ and $N$ when the costs of govern an $N+1$-sized city council is lower than the costs of govern an $N$-sized city council:

@@

This provides us our first empirical hypothesis to test: if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase.

At this stage, we need to solve for the costs of govern to determine whether this hypothesis is true or not. Below we show that it is true for the *partisan* and *hybrid* mechanisms, but it is not true for the *non-partisan* mechanism.

@@

In the \textit{non-partisan reversal} mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. To find the stationary sub-game perfect Nash equilibrium, suppose that there were $k-1$ rejections, and we are at the $k$-th stage in the game. A councilor accepts the proposer's offer if, and only if, to accept the offer is better than to wait until the next stage. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left( 1 - \dfrac{1}{N} \right)\left( \dfrac{1}{2} \right)\left[ x_i \right]
\]

In the left-hand side, we place the offer. In the right-hand side, there are two components. The first is the amount that the councilor $i$ gets when she is the proposer. It is equal to the budget in the next round minus the offers that she has to make to convince half of the councilors. The second part is the gains if when the councilor happens to reject the current received offer, but still receives an offer in the following round. Note that the offer the councilor makes as the proposer is the same as the offer that she wants to receive. This because the city councilors are exchangeable, and the solution is symmetric for all councilors receiving an offer (this means that we could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k)
\]

The proposer always offer the minimum possible to get the proposal approved. In the case, the offer at any given stage $k$ is going to be equal to $\underline{x}(k)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the costs of govern are equal to:

\[
C_G(N) \ = \ \dfrac{\delta R N}{2N + 1}
\]

\begin{proposition}
In the \textit{non-partisan reversal} mechanism, the costs of govern increase as the size of the legislature increases.
\end{proposition}

@@

This result is crucial to understand the role of the politics in our model. If the costs of govern increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In our case, we show that there is a \textit{political} reason for increasing the public goods provision.

@@

We now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p > 0$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a cost of $-p$. Moreover, as in the case of the non-partisan reversal, the mayor can choose to distribute $x_i$ private goods for a legislator $i$. In this case, when the policy is implemented, the utility of an aligned legislator is equals to $x_i + p$. When the legislator is not aligned with the mayor, her utility is equals to $x_i - p$. If the policy is not implemented, then we assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor always support the mayor's policy, regardless of receiving or not private goods. This because as $p > 0$, the councilor always receive positive benefits from the policy.\footnote{The mayor could even impose a cost to the city councilors aligned with the government. Although this seem unrealistic, it is mostly the case when government-aligned legislators are recognized to draft legislation for the mayor's office, or even take the popular fallout from a failed policy.} However, politicians not aligned with the mayor require a compensation when their votes are needed. In this case, if the mayor's proposal requires the support of an opposition politician, then the mayor compensates the councilor for the policy costs. In such case:

@@

Optimality dictates that the mayor offers $x_i = p$ to the opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of govern when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor has to compensate minus the amount needed to implement the policy, in this case, half of them:

@@

This result shows that when the chances of electing an aligned politician is sufficiently high, then the costs of govern decrease. As a response, the public goods provision (and also rents), increase. The rate distribution between these two vectors is determined by the marginal change in the utility from rents and the utility from reelection. In any case, the amount allocated for both increase.

@@

We can decompose these costs in two components. First, the costs in terms of rents. Second, the costs (or benefits) from political alignment. In this context, the governing costs depend on the mayor's offers for councilors in the mayor's coalition versus the mayor's offers for councilors in the opposition. The ex-ante chance of a opposition member be elected is equal to $1-\gamma$, and taking the weighted averages, these costs become:

@@

The conditions for the differences in governing costs be decrease when the  legislature size increases is:

@@

\begin{corollary}
In the \textit{hybrid reversal} mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

Also, the threshold $\underline{\gamma}$ for the probability of electing a politician aligned with the government is decreasing in both the political value of the mayors policy $p$ and the size of the legislature $N$. This means that higher political values and larger legislatures makes it easier to satisfy the electoral threshold.

@@

Our model provides two empirically testable hypotheses. Consider a city council comprised of $N$ legislators, that increase its council size to $N+1$. Then:

\begin{enumerate}
  \item[H1. ] The provision of public goods increase when the governing costs decrease with legislature size.
  \item[H2. ] The governing decrease with legislature size when the chance of electing a government aligned legislator is sufficiently high.
\end{enumerate}

In the paper, we show that the provision of education and healthcare improves in Brazil, and this results from the fact that the chances of electing a city councilor aligned with the mayor is of 91%, thanks to the rule of distributing seats adopted in the country.

@@

We use three information sources, either from Brazilian governmental agencies or available online. Table \ref{varsources} displays the primary sources and their respective URLs.

\begin{sidewaystable}[!htbp] \centering 
\caption{Data Sources} 
\label{varsources}
\scalebox{0.75}{
  \begin{tabular}{@{\extracolsep{5pt}}llll} 
  \\[-1.8ex]\hline\hline \\[-1.8ex] 
  Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{URL} \\ 
  \hline \\[-1.8ex] 
  & & & \\
  DataSUS & Brazilian Health Ministry Data Service & Collects data on Health Care & \url{http://www.datasus.gov.br} \\
  & & & \\
  IBGE & Brazilian Institute of Geography and Statistics & Collects data Geography, Economics and Demography & \url{http://www.ibge.gov.br} \\ 
  & & & \\
  INEP & Data Service of Ministry of Education & Collects data on performance of Education & \url{http://www.inep.gov.br} \\ 
  & & & \\
  InterLegis & Senate Legislative Data Service & Collect Data on Legislative in Brazil & \url{http://www.interlegis.leg.br} \\
  & & & \\
  IPEA & Brazilian Institute of Applied Economics & Collects data on the Economy & \url{http://www.ipeadata.gov.br} \\ 
  & & & \\
  MDS & Social Security Ministry & Collect data on Social Coverage and Effectiveness of Social Programs & \url{http://www.mds.gov.br} \\ 
  & & & \\
  TSE & Superior Electoral Tribunal & Collects data on elections results & \url{http://www.tse.jus.br} \\
  & & & \\
  \hline\hline \\[-1.8ex] 
  \end{tabular} 
}
\end{sidewaystable}

Now let us define each variable.

@@

We coded the legislation approved by the councilors in 64 municipalities that are 10 thousand inhabitants away from the council size thresholds. We classified the legislation in five categories:

\begin{itemize}
\item \textbf{Local Public Goods}: Legislation that provides a service targeted at the individual, group, or geographical level.
\item \textbf{Public Goods or General Legislation}: Legislation that provides a service for the entire municipality, or approves a law that does not discriminate among citizens.
\item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
\item \textbf{Others}: Legislation that is not classified as Local Public Goods, Public Goods, or Oversight. Usually honors or procedures.
\item \textbf{Education and Health Care}: Legislation about Education or Health Care provision.
\end{itemize}

@@

We run an online survey asking 174 councilors whether helping with [...service...] gives [none, few, some, or many] votes. The services were:

>>>>

@@

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch and has the prerogative of proposing a vector of policies voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors \textbf{x}.

The city council votes the mayor's proposal, and if they accept it, it is implemented. Otherwise, a reversal policy takes place. Reversal policy in this context means that the mayor's proposal failed to secure the majority of council votes. The decision process leaves the mayor's office and either immediately stops or is transferred to the council. The types of reversal policies are crucial for my argument, as it changes the relative strength of the council in their bargaining with the mayor. I investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan affiliation; a \textit{partisan} reversal mechanism, where the council is divided into supporters and non-supporters of the mayor; and finally, a \textit{hybrid} reversal mechanism, that combines partisan with non-partisan features. The main paper focuses on the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms clearly illustrates my argument that the partisan mechanism drives my results.

The mayor has to convince at least half of the city councilors to support her policy proposals. Similar to \citet{buenodemesquita2005logic}, I assume that the selectorate, i.e., the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, the minimum number of members who have to be convinced to implement the policy, equals half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor or the councilor recognized as the proposer \textbf{x}. I leave the definition of these transfers open, as this facilitates the relationship with several instruments that mayors can use to pressure city councilors. To illustrate, examples of transfers are granting electoral favors, spending in areas that the councilor has political or personal interests, giving portfolio and public jobs within the municipal bureaucracy, or even campaign contributions and bribes. The use of transfers affects the councilors' choices and generates a governing cost of $C_G$. The governing costs vary with the reversal mechanism and the legislature size.

Finally, after paying the govern costs, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision helps the mayor to get reelected. Rents are for direct mayor's consumption and do not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents $u(.)$ and the probability of reelection $\pi(.)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate.\footnote{In technical terms, $u'>0$ and $\pi'>0$; and $u''<0$ and $\pi''<0$.} The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. I assume that the benefits are high enough to rule out an equilibrium with only rents and transfers, which is equivalent to saying that mayors strongly prefer to be reelected. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

@@

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R > 0$ resources, and cannot run debts, meaning that the budget must be balanced. I also assume that there are enough resources for the mayor to govern.

Let $C_G$ be the costs of governing. The budget balance constraint requires the offers that the mayor makes to the city councilors to satisfy the following inequality:

@@

The expected utility for the city councilors depends on the reversal policy. As discussed, I explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to propose. The process repeats until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, I assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, I say that she belongs to the government coalition $G \subset \{1, 2, \cdots, N \}$. Otherwise, the city councilor belongs to the opposition $O = \{1, 2, \cdots, N \} \setminus G$. The mayor's policy generates a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends on the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor that should have received $p$ from the policy now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy and now receive zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent defying the partisan label.

Lastly, in the \textit{hybrid} reversal mechanism, I combine partisanship with legislature bargaining costs. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the non-partisan reversal mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. On the other hand, opposition members receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as the city council does not control the Executive when the two branches are separated, no public goods are provided ($p=0$). Moreover, in equilibrium, I assume that city councilors would prefer to receive transfers instead of providing public goods. This aligns with the empirical evidence in most developing economies.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented, and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, I propose three different reversal policies. The details are in the following sections.
\end{enumerate}

The solution concept I use in this model is the Sub-game Perfect Nash Equilibrium. A Sub-game Perfect Nash Equilibrium requires that the strategies played are a Nash Equilibrium in each of the subgames of the game. The solution is found using backward induction. In the model with infinitely repeated proposals, I extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium requires that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should take the same offer at time $k$. Stationarity provides the intending to find a proposal that would be accepted at any stage in the game. Applying this reasoning, I can find a sequence of offers for each point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a no delay in policy implementation.

\begin{definition}
The equilibrium for the game consists of a sequence of strategic policy vectors $(g^*, r^*, x^*)_k$, indexed in $k \in \{0, 1, \cdots, \infty\}$, such that, at every stage of the game, the mayor and the councilors have no profitable deviations from the equilibrium.
\end{definition}

Moreover, I assume that the players play no dominated strategies, even when there is less than sufficient support for implementing the policy.

@@

Solving the game by backward induction requires me to start with the councilor's strategies. However, for exposure convenience, assume that I solved the game for the councilors and found the equilibrium cost of governing $C_G(N)$ when there are $N$ city councilors. In the next section, I define $C_G(N)$ for each mechanism using backward induction, and the definitions are well-behaved. In this section, I use this assumption to simplify the exposition.

I now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as necessary for the public goods and extract the remaining resources as political rents. The mayor's objective is to maximize her expected utility, subject to the municipal budget constraint.

@@

In equilibrium, the budget binds, and the provision of the public goods maximizes:

@@

The first-order condition for an optimal public goods provision makes the marginal benefits to increase rents, in this case, marginal costs of providing the public goods, equals to the marginal benefits of reelection:

@@

\begin{proposition}
The provision of public goods increases with legislature size when the costs of governing decrease with legislature size.
\end{proposition}

\begin{proof}
To prove this result, I need to show that the mayor's expected utility satisfies the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, I cannot take the cross-partial derivative on $N$ or even use the implicit function theorem. Instead, I use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

The mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1 > N$,\footnote{Moreover, the requirement that the decision and type space is a complete lattice, is satisfied for $\mathbb{R} \times \mathbb{N}$.} I have:

@@

After substituting, I have:

@@

Multiplying both sides by $\dfrac{1}{g - g'}$ and taking the limit when $g \rightarrow g'$, I have:

@@

It is now clear to see that the function satisfies the increasing differences in $g$ and $N$ when the governing costs of a $N+1$-sized city council are lower than the governing costs of a $N$-sized city council:

@@

This provides the first empirical hypothesis to test: \textit{if the governing costs decrease when the legislature increases, then the public goods provision increase when the legislature size increase}.

At this stage, I need to solve for the existence of these governing costs. Below I show that this proposition is valid for the *partisan* and *hybrid* mechanisms, but it is not suitable for the *non-partisan* mechanism.

@@

In the \textit{non-partisan reversal} mechanism, I assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. To find the stationary sub-game perfect Nash equilibrium, suppose that there were $k-1$ rejections, and the game is at the $k$-th stage. A councilor accepts the proposer's offer if, and only if, to get the offer at $k$ is better than to wait until the next stage $k+1$. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left( 1 - \dfrac{1}{N} \right)\left[\dfrac{1}{2} x_i \right]
\]

On the left-hand side, I place the offer. On the right-hand side, there are two components. The first is the amount that the councilor $i$ gets when she is the proposer. It is equal to the chance that she is recognized as the proposer times the budget in the next round minus the offers she has to make to convince half of the councilors. The second part is the gains if the councilor rejects the current received offer but still gets an offer in the following round. It is equal to the chance that she is not recognized as the proposer times the chance that she receives a transfer times the transfer amount. Note that the councilor's offer as the proposer is the same as the offer she wants to receive. This is because the city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, I could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k, N)
\]

The proposer always offers the minimum required to get the proposal approved. In this case, the offer at any given stage $k$ is going to be equal to $\underline{x}(k, N)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the governing costs for the mayor are:

\[
C_G(N) \ = \ \dfrac{N}{2}\left[\dfrac{2 \delta R }{2N + 1}\right] \ = \ \dfrac{\delta R N}{2N + 1}
\]

\begin{proposition}
In the \textit{non-partisan reversal} mechanism, the governing costs always increase when the size of the legislature increases.
\end{proposition}

@@

This result is crucial to understand the role of politics in my model. If the governing costs increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In my case, I show that there is a \textit{political} reason for increasing the public goods provision.

@@

I now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p > 0$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a cost of $-p$. Moreover, as in the non-partisan reversal, the mayor can choose to distribute $x_i$ private goods for a legislator $i$. In this case, when the policy is implemented, the aligned legislator's utility equals $x_i + p$. When the legislator is not aligned with the mayor, her utility equals $x_i - p$. If the policy is not implemented, then I assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor always supports the mayor's policy, regardless of receiving private goods. This is because as $p > 0$, the councilor always receives positive benefits from the policy.\footnote{The mayor could even impose a cost to the city councilors aligned with the government. Although this seems unrealistic, it is mostly the case when government-aligned legislators are recognized to draft legislation for the mayor's office or even take the fallout from a failed policy.} However, politicians not aligned with the mayor require compensation when their votes are needed. In this case, if the mayor's proposal requires the support of an opposition politician, then the mayor compensates the councilor for the policy costs. In such case:

@@

Optimality dictates that the mayor offers $x_i = p$ to the opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of governing when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor has to compensate minus the amount needed to implement the policy, in this case, half of them:

@@

This result shows that when the chances of electing an aligned politician are sufficiently high, the government costs decrease. As a response, the public goods provision (and also rents) increase. The rate distribution between these two vectors is determined by the marginal change in the utility from rents and the utility from reelection. In any case, the amount allocated for both increases.

@@

I can decompose these costs into two components—first, the costs in terms of rents. Second, the costs (or benefits) from political alignment. In this context, the governing costs depend on the mayor's offers for councilors in the mayor's coalition versus the mayor's proposals for councilors in the opposition. The ex-ante chance of an opposition member be elected equal to $1-\gamma$, and taking the weighted averages, these costs become:

@@

The conditions for the differences in governing costs decrease when the legislature size increases are:

@@

\begin{corollary}
In the \textit{hybrid reversal} mechanism, when the chances of electing a city councilor aligned with the mayor are sufficiently high, then increasing council size lowers the bargaining costs for mayors.
\end{corollary}

The threshold $\underline{\gamma}$ for the probability of electing a politician aligned with the government is decreasing in the value of the mayor's policy $p$ and the legislature size $N$. This means that higher political values and larger legislatures make it easier to satisfy the electoral threshold.

@@

Our model provides two empirically testable hypotheses. Consider a city council comprised of $N$ legislators that increase its council size to $N+1$. Then:

\begin{enumerate}
  \item[H1. ] The governing costs decrease with legislature size when the chance of electing a government-aligned legislator is sufficiently high.
  \item[H2. ] The provision of public goods increase when the governing costs decrease with legislature size.
\end{enumerate}

In the paper, I show that the provision of education and healthcare improves in Brazil. This results from the fact that the chances of electing a city councilor aligned with the mayor are 91%.

@@

I use three information sources, either from Brazilian governmental agencies or available online. Table \ref{varsources} displays the primary sources and their respective URLs.

\begin{sidewaystable}[!htbp] \centering 
\caption{Data Sources} 
\label{varsources}
\scalebox{0.8}{
  \begin{tabular}{@{\extracolsep{5pt}}llll} 
  \\[-1.8ex]\hline\hline \\[-1.8ex] 
  Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{URL} \\ 
  \hline \\[-1.8ex] 
  & & & \\
  DataSUS & Brazilian Health Ministry Data Service & Collects data on Health Care & \url{http://www.datasus.gov.br} \\
  & & & \\
  IBGE & Brazilian Institute of Geography and Statistics & Collects data Geography, Economics and Demography & \url{http://www.ibge.gov.br} \\ 
  & & & \\
  INEP & Data Service of Ministry of Education & Collects data on performance of Education & \url{http://www.inep.gov.br} \\ 
  & & & \\
  InterLegis & Senate Legislative Data Service & Collect Data on Legislative in Brazil & \url{http://www.interlegis.leg.br} \\
  & & & \\
  IPEA & Brazilian Institute of Applied Economics & Collects data on the Economy & \url{http://www.ipeadata.gov.br} \\ 
  & & & \\
  MDS & Social Security Ministry & Collect data on Social Coverage and Effectiveness of Social Programs & \url{http://www.mds.gov.br} \\ 
  & & & \\
  TSE & Superior Electoral Tribunal & Collects data on elections results & \url{http://www.tse.jus.br} \\
  & & & \\
  \hline\hline \\[-1.8ex] 
  \end{tabular} 
}
\end{sidewaystable}

@@

I coded the legislation approved by the councilors in 63 municipalities within ten thousand inhabitants away from the council size thresholds. I classified the legislation into five categories:

\begin{itemize}
\item \textbf{Public Goods}: Legislation that provides public goods or services.
\item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
\item \textbf{Others}: Legislation that is not classified as Local Public Goods, Public Goods, or Oversight. Usually honors or procedures.
\item \textbf{Education and Health Care}: Legislation about Education or Health Care provision.
\end{itemize}

@@

I ran an online survey asking 174 councilors, among other questions, about their representation duties, which mayors used services to improve the mayor's coalition support. The services were:

\begin{itemize}
\item \textbf{Councilors Job Appointments}: The frequency that mayors appoint people allied to city councilors to bureaucratic jobs (source: survey).
\item \textbf{Councilors Demands}: The frequency that mayors satisfy an allied councilors' demand (e.g., build a health clinic in a councilor's neighborhood) (source: survey).
\item \textbf{Councilors Personal Requests}: The frequency that mayors satisfy an allied councilors' requests (e.g., hire an ally in the bureaucracy) (source: survey).
\item \textbf{Councilors Legislation}: The frequency that mayors support councilors' legislative agenda (source: survey).
\item \textbf{Councilors Constructions}: The frequency that mayors support councilors' requests for constructions (e.g., fix potholes in the councilor's neighborhoods) (source: survey).
\end{itemize}

This Appendix presents other questions I asked in the survey. They are meant to improve my knowledge about the city councilors' perceptions of their municipality roles. I omitted these questions in the main paper because they are not of general interest to my argument. The descriptive statistics of the variables used in the main article follow in Table \ref{tabdescrvarspaper}.
